---
title: "Final Project R"
author: "Surbhi Rathore"
date: "4/26/2021"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, echo = FALSE, warning=FALSE}
suppressPackageStartupMessages({library(tidyr)
library(dplyr)
library(tinytex)
library(ggplot2)
library(ggplot2)
library(tmap)
library(deSolve)
library(date)
library(moderndive)
library(modelr)
library(dplyr)
library(sf)
require(tree)
library(MASS)})

#Reading csv files for accessing datasets: My Analyses is based on Covid 19 data, hence I am using the confirmedraw, deathraw and recoveredraw datasets from Johns Hopkins Github data

confirmedraw <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")

deathsraw <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")

recoveredraw <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")


```
##DATASET
Data sources: Coronavirus COVID-19 Global Cases by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University. 

```{r, echo = FALSE, warning=FALSE}
# DATA CLEANING: To create country level and global combined data and Convert each data set from wide to long AND aggregate at country level.

# Refining country level confirmed cases and aligning wide to long
confirmed <- confirmedraw %>% gather(key="date", value="confirmed", -c(Country.Region, Province.State, Lat, Long)) %>% group_by(Country.Region, date) %>% summarise(confirmed=sum(confirmed, na.rm = TRUE))

# Refining country level death cases and aligning wide to long
deaths <- deathsraw %>% gather(key="date", value="deaths", -c(Country.Region, Province.State, Lat, Long)) %>% group_by(Country.Region, date) %>% summarise(deaths=sum(deaths, na.rm = TRUE))

# Refining country level recovered cases and aligning wide to long
recovered <- recoveredraw %>% gather(key="date", value="recovered", -c(Country.Region, Province.State, Lat, Long)) %>% group_by(Country.Region, date) %>% summarise(recovered=sum(recovered, na.rm = TRUE))

# Combining all the three refined datasets using full join
country <- full_join(confirmed, deaths) %>% full_join(recovered)

#str(country) # check date character

# Date variable, Fixing date variable and converting from character to date
country$date <- country$date %>% sub("X", "", .) %>% as.Date("%m.%d.%y")

#str(country) # check date Date

#Plot to show the confirmed cases all over the country
ggplot(country, aes(x=date, y=confirmed, col = 'blue')) + geom_bar(stat="identity", width=0.1) +
  theme_classic() +
  labs(title = "Covid-19 Confirmed Cases with respect to countries", x= "Date", y= " confirmed cases") +
  theme(plot.title = element_text(hjust = 0.5))

#Plot to show the death cases all over the country
ggplot(country, aes(x=date, y=deaths, col = 'black')) + geom_bar(stat="identity", width=0.1) +
  theme_classic() +
  labs(title = "Deaths due to Covid-19 with respect to countries", x= "Date", y= " death cases") +
  theme(plot.title = element_text(hjust = 0.5))

#Plot to show the recovered cases all over the country
ggplot(country, aes(x=date, y=recovered ,col = 'red')) + geom_bar(stat="identity", width=0.1) +
  theme_classic() +
  labs(title = "Covid-19 recovered Cases with respect to countries", x= "Date", y= " recovered cases") +
  theme(plot.title = element_text(hjust = 0.5))


```
##Coronavirus disease 2019 (COVID-19), 
also known as the coronavirus or COVID, is a contagious disease caused by severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2). The first known case was identified in Wuhan, China, in December 2019. The disease has since spread worldwide, leading to an ongoing pandemic. Thus, It became more important to analyze the disease and understand the statistics to spread awareness among people. Today when there is no immunity against the disease, we can probably understand the statistics of how precarious the disease is and how much death it has caused and can cause in future.

##Observations:
First plot of confirmed cases shows that the spread took over more that 140 million people till January 2021. The wave stated in the month of December 2019 and led to several symptoms that existed for 14 or more days followed by testing results.

Second plot shows the number of death cases due to covid 19 virus. the numbers started ranging from zero to 100, 1000 and reached up to a million by july 2020. currently number of deaths have reached upto 3 million.

The third plot shows the recovered cases. It account for people who were infected but recovered and the current number is 75 million, 2021. 


```{r,echo = FALSE, warning=FALSE}
#Further filteration of data frame
country <- filter(country,  confirmed > 0, deaths > 0, recovered > 0)

## Creating new variable: number of days
country <- country %>% group_by(Country.Region) %>% mutate(cumconfirmed=cumsum(confirmed), days = date - min(date)  + 1)
#names(country)

##grouping cumulative confirmed case and associated region
country_confirmed <- country %>% group_by(Country.Region) %>% mutate(cumconfirmed=cumsum(confirmed))%>% summarize(cumconfirmed=sum(cumconfirmed, na.rm = TRUE))

##returning unique values from country.region column to run analysis between countries.
#unique(country[c("Country.Region")])

## Aggregate at world level
world <- country %>% group_by(date,Country.Region) %>% summarize(confirmed=sum(confirmed, na.rm = TRUE), cumconfirmed=sum(cumconfirmed, na.rm = TRUE), deaths=sum(deaths, na.rm = TRUE), recovered=sum(recovered, na.rm = TRUE))%>% mutate(days = date - min(date) + 1)


## Extract specific country: India and USA
india <- country %>% filter(Country.Region=="India")
usa <- country %>% filter(Country.Region=="US")

##Summary statistics with respect to confirmed, deaths, recovered cases and cumulative confirmed covid case 
#summary(country)
#by(country$confirmed, country$Country.Region, summary)
#by(country$deaths, country$Country.Region, summary)
#by(country$recovered, country$Country.Region, summary)

```


```{r,echo = FALSE, warning=FALSE}
#summary(world)
summary(india)
summary(usa)
```

##Summary of India
Mean of the confirmed cases for India was recorded as 5310060 by september 2020 while the mean of death cases was recorded as 77553 followed by recovered cases around 4850399.

##Summary of USA
Mean of the confirmed cases for USA was recorded as 10770420 by september 2020 while the mean of death cases was recorded as 226267 followed by recovered cases around 1078030.

Indian population and American population has huge difference, no conclusion can be drawn with the analysis but we can get and an idea of the current numbers.

```{r, echo = FALSE, warning=FALSE}
#Plots

# plot of World confirmed cases
ggplot(world, aes(x=date, y=confirmed)) + geom_bar(stat="identity", width=0.1) +
  theme_classic() +
  labs(title = "Covid-19 Global Confirmed Cases", x= "Date", y= "Daily confirmed cases") +
  theme(plot.title = element_text(hjust = 0.5))

# India confirmed cases
ggplot(india, aes(x=days, y=confirmed)) + geom_bar(stat="identity", width=0.1) +
  theme_classic() +
  labs(title = "Covid-19 Confirmed Cases in India", x= "Increased cases over Days", y= "Confirmed cases") +
  theme(plot.title = element_text(hjust = 0.5))

# USA confirmed cases
ggplot(usa, aes(x=days, y=confirmed)) + geom_bar(stat="identity", width=0.1) +
  theme_classic() +
  labs(title = "Covid-19 Confirmed Cases in USA", x= "Increased cases over Days", y= "Confirmed cases") +
  theme(plot.title = element_text(hjust = 0.5))

# World confirmed, deaths and recovered
#str(world)
world %>% gather("Type", "Cases", -c(date, days, Country.Region)) %>%
  ggplot(aes(x=date, y=Cases, colour=Type)) + geom_bar(stat="identity", width=0.2, fill="white") +
  geom_line()+
  geom_point()+
  theme_classic() +
  labs(title = "Covid-19 Global Cases", x= "Date", y= "Daily cases") +
  theme(plot.title = element_text(hjust = 0.5))


```

##Observations:

Looking at the plots for data set globally, we can make following relevant observations:

*The sum of data is growing with time. This lets us view the total contribution so far of a given measure against time.

*Number figures also shows us that deaths are increasing as well, which notifies of how harmful effects of the virus is.

*Though we have recovered case numbers yet it does not rule out the death numbers and demands people to follow safety measures to cope up with the pandemic.


```{r, include=FALSE,echo = FALSE, warning=FALSE}


# Map
countrytotal <- country %>% group_by(Country.Region) %>% summarize(cumconfirmed=sum(confirmed), cumdeaths=sum(deaths), cumrecovered=sum(recovered))

#Basemap from package tmap
data(World)
class(World)

# Combine basemap data to covid data
countrytotal$Country.Region[!countrytotal$Country.Region %in% World$name]

list <- which(!countrytotal$Country.Region %in% World$name)

countrytotal$country <- as.character(countrytotal$Country.Region)

countrytotal$country[list] <-
  c("Andorra", "Antigua and Barbuda", "Bahrain",
    "Barbados", "Bosnia and Herz.", "Myanmar",
    "Cape Verde", "Central African Rep.", "Congo",
    "Dem. Rep. Congo", "Czech Rep.", "Diamond Princess",
    "Dominica", "Dominican Rep.", "Eq. Guinea",
    "Swaziland", "Grenada", "Holy See",
    "Korea", "Lao PDR", "Liechtenstein",
    "Maldives", "Malta", "Mauritius",
    "Monaco", "MS Zaandam", "Macedonia",
    "Saint Kitts and Nevis", "Saint Lucia", "Saint Vincent and the Grenadines",
    "San Marino", "Sao Tome and Principe", "Seychelles",
    "Singapore", "S. Sudan", "Taiwan",
    "United States", "Palestine", "W. Sahara", "India")

countrytotal$Country.Region[!countrytotal$country %in% World$name]

World$country <- World$name

worldmap <- left_join(World, countrytotal, by="country")
worldmap$cumconfirmed[is.na(worldmap$cumconfirmed)] <- 0



```


```{r, echo=FALSE,warning=FALSE}

# Confirmed cases by country for select countries with counts in log10 scale
countryselection <- country %>% filter(Country.Region==c("US", "Italy", "India","China", "France", "United Kingdom", "Germany"))
ggplot(countryselection, aes(x=days, y=confirmed, colour=Country.Region)) + geom_line(size=1) +
  theme_classic() +
  labs(title = "Covid-19 Confirmed Cases by different Countries", x= "Days", y= "Daily confirmed cases (log scale)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(trans="log10")

# Matrix of line graphs of confirmed, deaths and recovered for select countries in log10 scale
#str(countryselection)
countryselection %>% gather("Type", "Cases", -c(date, days, Country.Region)) %>%
  ggplot(aes(x=days, y=Cases, colour=Country.Region)) + geom_line(size=1) +
  theme_classic() +
  labs(title = "Covid-19 Cases by Country", x= "Days", y= "Daily cases (log scale)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(trans="log10") +
  facet_grid(rows=vars(Type))

# Map
ggplot(data = worldmap) + geom_sf(aes(fill=cumconfirmed), color="black") +
  ggtitle("World Map of Confirmed Covid Cases",
          subtitle="Total Cases on April 20, 2020") +
  theme_bw()
```

*First Graph shows us the numbers from different countries like: China, France, Germany, India, Italy, UK, and US.*

Observations:

The disease is wide spreading, though it started from china, but china is stabilizing the numbers by taking strict measures. whereas countries like India and US still have no control over the disease and is taking over uninfected people thus increasing the numbers. Other countries, France, Italy UK, Germany showing increasing number but little less than numbers of India and US.

*Second Graph shows us collaborative reports for confirmed, deaths, recovered and cumulative confirmed cases for countries specified for graph 1.*

Observations:

The most promising results in the graph for all the 3 cases is for China, even though the country led to spread of this deadly virus, it played better role in controlling the infection and bending down the death rates as well.
Another promising result is the recovered cases from India, even though the cumulative confirmed cases are higher but some how the virus is leading to less deaths compared to other countries. Reason may be better immunity or health care infrastructure (It's just speculated idea). Hence India shows highest number of recovered cases.
In addition, US shows least number in the recovered cases, where as the confirmed cases are increasing by each passing days, even the death cases have higher number compared to other countries.

*Third Graph is a World map which shows cumulative confirmed cases all over the world*

Observations:

It shows the numbers in shades of blue and has recognized least to increasing number of cases.
Brazil covers the lightest shade of blue with most of the cumulative confirmed cases.


```{r, echo=FALSE,warning=FALSE}
#Notes for future references:
#1.I tried SIR model on Country.Region= India and US but somehow it does not show me results as required.
#2.I tried different values for gamma and beta but for any value less in 0.5 for gamma and beta the model does not work also, it gives me 0 R values with gamma= 0.5, which is not expected. As the number of susceptible cases goes high following the infected cases the recovered cases should show some number.
#3. I plan to work on optimization of the model for better results.

## MODEL
# Analysis on region Ontario
ontario <- confirmedraw %>% filter(Province.State=="Ontario") %>% gather(key="date", value="confirmed", -c(Country.Region, Province.State, Lat, Long)) %>%  mutate(cumconfirmed=cumsum(confirmed))
ontario$date <- ontario$date %>% sub("X", "", .) %>% as.Date("%m.%d.%y")

# SIR FUNCTION - SIR function gives us analysis on susceptible, infected and recovered group of people as predictions for future.
SIR <- function(time, state, parameters) {
  par <- as.list(c(state, parameters))
  with(par, {
    dS <- -beta * I * S/N
    dI <- beta * I * S/N - gamma * I
    dR <- gamma * I
    list(c(dS, dI, dR))
  })
}

# CREATE A VECTOR OF DAILY CUMULATIVE INCIDENCE NUMBERS OF CANADA FROM START DATE
infected <- ontario %>% filter(confirmed>0) %>% pull(cumconfirmed)

# Create an incrementing Day vector the same length as our cases vector
day <- 1:(length(infected))
N <- 14446515

# now specify initial values for S, I and R
init <- c(S = N - infected[1], I = infected[1], R = 0)

#Residual Sum of Squares-measures the level of variance in the error term, or residuals, of a regression model
RSS <- function(parameters) {
  names(parameters) <- c("beta", "gamma")
  out <- ode(y = init, times = day, func = SIR, parms = parameters)
  fit <- out[, 3]
  sum((infected - fit)^2)
}

# now find the values of beta and gamma that give the smallest RSS, which represents the best fit to the data.
# using method L-BFGS-B to constrain the interval 0 to 1.0
optimization <- optim(c(0.5, 0.5), RSS, method = "L-BFGS-B", lower = c(0,0), upper = c(1, 1))

#check for convergence
optimization$message

# Optimization Parameters
opt_par <- setNames(optimization$par, c("beta", "gamma"))
opt_par

#Reproduction Number
R0 <- opt_par[1]/opt_par[2]
R0

# PREDICTION - time in days for predictions
t <- 1:150

# get the fitted values from our SIR model
fitted <- data.frame(ode(y = init, times = t, func = SIR, parms = opt_par))

# add a Date column and join the observed incidence data
startdate <- "2020-01-22" # first occurrence recorded date
fitted <- fitted %>%
  mutate(date = as.Date(startdate) + t - 1, province = "Ontario") %>%
  left_join(ontario, by = c("date"="date"))

#Removing unwanted columns
fitted <- subset(fitted, select = -c(Province.State, province, time, Country.Region, Lat, Long, confirmed))


# plot the data
ggplot(fitted, aes(x = date)) +
  geom_line(aes(y = I), colour = "red") +
  geom_point(aes(y = cumconfirmed), colour = "orange")
labs(y = "Cumulative incidence", x="Date",
     title = "COVID-19 fitted vs observed cumulative incidence, Ontario",
     subtitle = "(red=fitted incidence from SIR model, orange=observed incidence)")

# plot the data
ggplot(fitted, aes(x = date)) +
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = S), colour = "black") +
  geom_line(aes(y = R), colour = "green") +
  geom_point(aes(y = cumconfirmed), colour = "orange") +
  scale_y_continuous(labels = scales::comma) +
  labs(y = "Persons", title = "COVID-19 fitted vs observed cumulative incidence, Ontario province") +
  scale_colour_manual(name = "",
                      values = c(red = "red", black = "black", green = "green", orange = "orange"),
                      labels = c("Susceptible", "Recovered", "Observed incidence", "Infectious")) +
  scale_y_continuous(trans="log10")

```
*For the first graph, I plotted infected(red line) along with cumulative confirmed cases(orange line)*

In general infection leads to confirmed cases hence the number should have gone higher for infected as well, but my model does not perform better in fetching results for infected numbers(red line). I plan on improving results with better approach.

*Second graph shows the SIR graph along with cumulative confirmed numbers*
The graph does show rising infection numbers but shows flat reading for both susceptible and recovered, which mean the infection is rising and the whole population is susceptible but there is no recovery rate, this output is not acceptable, considering the current scenario.
This prediction model is basically not accurate as it should show figures and numbers based on current scenario, where people are getting infected and few have been recovered too along with few deaths.

```{r, echo = FALSE, warning=FALSE}
#linear regression model on worlds data 
lmc <- lm(cumconfirmed ~ pop_est_dens, data = worldmap) #confirmed cases can be related to the estimated population density
summary(lmc)

get_regression_table(lmc)

par(mfrow = c(2, 2))
plot(lmc)

#Multiple regression model on worlds data
# population density and total population along with economical condition can play major role in spread of the virus
lmcm<-lm(cumconfirmed ~ pop_est_dens+ pop_est + economy, data = worldmap)
summary(lmcm)

get_regression_table(lmcm)

par(mfrow = c(2, 2))
plot(lmcm)

```
*Linear Model*
The Pr(>t) acronym found in the model output relates to the probability of observing any value equal or larger than t. A small p-value indicates that it is unlikely we will observe a relationship between the predictor (pop_est_dens) and response (cumconfirmed) variables due to chance. Typically, a p-value of 5% or less is a good cut-off point. 
In our model, the p-values are very close to zero. Three stars (or asterisks) represent a highly significant p-value. Consequently, a small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between pop_est_dens and cumconfirmed cases.

*Multiple Linear Regression Model*
In multiple linear regression, the R2 represents the correlation coefficient between the observed values of the cumconfirmed and the fitted i.e.pop_est_dens + pop_est + economy values of cumconfirmed, For this reason, the value of R will always be positive and will range from zero to one.
A problem with the R2, is that, it will always increase when more variables are added to the model, even if those variables are only weakly associated with the response. A solution is to adjust the R2 by taking into account the number of predictor variables.

The adjustment in the “Adjusted R Square” value in the summary output is a correction for the number of predictor variables included in the prediction model. The lower the RSE, the more accurate the model, In our case RSE is 0.4902.

A small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between pop_est_dens + pop_est + economy and cumconfirmed cases.

*Plot1*
The results show a strong evidence that estimated population density is the major influencing cumulative confirmed cases.

Shown residual plots also adds strength for the same.
There are few outliers found while performing the linear fit.

Approximately 90+% of variation in value variable can be explained by this model with these two independent variables (pop_est_dens and cumconfirmed).
Very low P-value also strengthens the assumption.
The residual standard error also shows there is not much distance between our observed value from the predicted Value.
We can see from the plots that there exist leverage points but, they are not much influential.

*Plot2*
The results show a strong evidence that estimated population density, total population estimate and economy collectively are the majorly influencing cumulative confirmed cases.



```{r, echo = FALSE, warning=FALSE}
#removing unwanted columns from existing worldmap table for further analysis
worldmap_a <- subset(worldmap, select = -c(iso_a3, name, sovereignt,footprint,inequality,HPI,Country.Region,geometry , well_being))

#removing na values from columns
worldmap_a<-worldmap_a %>% 
  filter(!is.na(continent))%>% 
  filter(!is.na(pop_est))%>% 
  filter(!is.na(pop_est_dens))%>% 
  filter(!is.na(economy))%>% 
  filter(!is.na(income_grp))%>% 
  filter(!is.na(gdp_cap_est))%>% 
  filter(!is.na(life_exp))%>% 
  filter(!is.na(country))%>% 
  filter(!is.na(cumconfirmed))%>% 
  filter(!is.na(cumdeaths))%>% 
  filter(!is.na(cumrecovered))


```

```{r, echo = FALSE, warning=FALSE}

#plotting the economy and confirmed cumulation cases to understand the effect in better way
ggplot(worldmap_a, aes(economy, cumconfirmed, color= economy))+
  geom_point()+
  geom_line()+
  geom_smooth(method = "lm",se=FALSE) #lm is linear model

#generating the result from lm method on world data considering the economy
model_worldmap_a<-lm(cumconfirmed ~ economy, data = worldmap_a)
#summary(model_worldmap_a)

get_regression_table(model_worldmap_a) #returns a neat table

#generating the result from lm method on world data considering population density estimated
model_worldmap_b<-lm(cumconfirmed ~ pop_est_dens, data= worldmap_a)
#summary(model_worldmap_b)
get_regression_table(model_worldmap_b) #returns a neat table


```

*Plot x: economy distribution and y: cumulative confirmed cases*

Well, the plot does not explains the relation between the spread of virus among people belonging to developing or emerging economy.
It shows variable results. We would agree with confirmed numbers shown for emerging region: MIKT for probable reasons like, they may not have better access to infrastructures or may have to meet people in order to get their work done which would have led to transmission of virus or many more.
But if we consider, Least developed region, they have the least number of cumulative confirmed cases, which is great considering their economical conditions.
In contrast to above note, even developed region has second highest number of cases.

This this graph does not conclude any thing in particular.

*Linear Model, Predictor- Economy*

To understand the relationship better, we did explore the linear model with economy as predictor and cumconfirmed as target, and to our surprise the model does work better. the P value and R squared both are close to 0, which explains the assumption.

*Linear Model, Predictor- population density estimate*

I had an inclination towards this particular model, as I believed the spread will increase if people assemble in closed area or are around others(Just an idea which is far away from the concept of population density but does explains the concept of density in a closed area). And this model explains it all, it gives us p-value: 0.2883 and Adjusted R-squared:  0.001011. It does explain that our model fits with the data provided and spread of the disease can be related to higher population density estimated.

```{r, echo = FALSE, warning=FALSE}
#adding  residuals, how much is calculated cases different from what is expected 
#based on calculating the mean of cases, residual is higher or lower than mean
worldmap_b<-worldmap_a%>% add_residuals(model_worldmap_a)

#plotting residuals and the cases confirmed numbers  
ggplot(worldmap_b, aes(cumconfirmed,resid, color= economy))+
  geom_ref_line(h=0)+
  geom_line()

#generating a linear model of residual with respect to different countries
model_temp_country<- lm(resid~ country, data =  worldmap_b)

#summary(model_temp_country)

```
Residuals: The basic idea behind finding residuals is that we can get the difference between predicted value and the measured values and find the line that fits our data set.


```{r, echo = FALSE, warning=FALSE}

# Dropping the geometry column for further analysis
worldmap_b_no_geometry <- st_set_geometry(worldmap_b, NULL)

# Confirm the geometry column has been dropped
#names(worldmap_b_no_geometry)

```

```{r, echo = FALSE, warning=FALSE}

set.seed(18)

#Basic Regression Trees
## Create the Training and Test dataset
#We first split the data in half.

idx = sample(1:nrow(worldmap_b_no_geometry), nrow(worldmap_b_no_geometry)/2)

worldmap_b_no_geometry_trn <- worldmap_b_no_geometry[idx,]

worldmap_b_no_geometry_tst <- worldmap_b_no_geometry[-idx,]

#Then fit an unpruned regression tree to the training data.
tree.worldmap_b_no_geometry <- tree(cumconfirmed ~ ., data = worldmap_b_no_geometry_trn)

summary(tree.worldmap_b_no_geometry)

plot(tree.worldmap_b_no_geometry)
text(tree.worldmap_b_no_geometry, pretty = 0)
title(main = "Unpruned Classification Tree")
```
##Regression Trees analysis
Performed another type of statistical model called Regression Trees, because it has a good graphical
representation.
*Basic Regression trees Covid 19 data set.
*The regression tree building methodology allows input variables to be a mixture of continuous and categorical variables.
*A decision tree is generated when each decision node in the tree contains a test on some input variable's value.
*The terminal nodes of the tree contain the predicted output variable values.

As with classification trees, we can use cross-validation to select a good pruning of the tree.

```{r, echo = FALSE, warning=FALSE}

set.seed(18)
worldmap_b_no_geometry_cv <- cv.tree(tree.worldmap_b_no_geometry)

plot(worldmap_b_no_geometry_cv$size, sqrt(worldmap_b_no_geometry_cv$dev / nrow(worldmap_b_no_geometry_trn)), type = "b",
     xlab = "Tree Size", ylab = "CV-RMSE")

```

*Cross-validation refers a methods for measuring the performance of a given predictive model on new test data sets.

The basic idea, behind cross-validation techniques, consists of dividing the data into two sets:
1.The training set, used to train (i.e. build) the model;
2. The testing set (or validation set), used to test (i.e. validate) the model by estimating the prediction error.


While the tree of size 380000000 does have the lowest RMSE, we’ll prune to perform just as well. The pruned tree is, as expected, smaller and easier to interpret.

```{r, echo = FALSE, warning=FALSE}

# pruning the tree, using cross validation

worldmap_b_no_geometry_prune<- prune.tree(tree.worldmap_b_no_geometry, best = 150000000)

summary(worldmap_b_no_geometry_prune)

plot(worldmap_b_no_geometry_prune)
text(worldmap_b_no_geometry_prune, pretty = 0)
title(main = "Pruned Regression Tree")

```
Regressing tree does not provide us better model view to understand the distribution of data, let’s compare this regression tree to an additive linear model and use RMSE as our metric.

```{r, echo = FALSE, warning=FALSE}
#Let’s compare this regression tree to an additive linear model and use RMSE as our metric.

rmse = function(actual, predicted) {
  sqrt(mean((actual - predicted) ^ 2))
}
```

We obtain predictions on the train and test sets from the pruned tree. We also plot actual vs predicted. This plot may look odd. We’ll compare it to a plot for linear regression below.

```{r, echo = FALSE, warning=FALSE}
# training RMSE two ways
#sqrt(summary(worldmap_b_no_geometry_prune)$dev / nrow(worldmap_b_no_geometry_trn)) # 302320691

worldmap_b_no_geometry_trn_pred = predict(worldmap_b_no_geometry_prune, newdata = worldmap_b_no_geometry_trn)

#rmse(worldmap_b_no_geometry_trn_pred, worldmap_b_no_geometry_trn$cumconfirmed) #302320691

# test RMSE
worldmap_b_no_geometry_tst_pred = predict(worldmap_b_no_geometry_prune, newdata = worldmap_b_no_geometry_tst)

#rmse(worldmap_b_no_geometry_tst_pred, worldmap_b_no_geometry_tst$cumconfirmed)#207699947



plot(worldmap_b_no_geometry_tst_pred, worldmap_b_no_geometry_tst$cumconfirmed, xlab = "Predicted", ylab = "Actual")
abline(0, 1)


```

Here, using an additive linear regression the actual vs predicted looks much more like what we are used to.

```{r, echo = FALSE, warning=FALSE}
worldmap_lm = lm(cumconfirmed ~ pop_est+pop_est_dens+economy, data = worldmap_b_no_geometry_trn)

worldmap_lm_pred = predict(worldmap_lm, newdata = worldmap_b_no_geometry_tst)

plot(worldmap_lm_pred, worldmap_b_no_geometry_tst$cumconfirmed, xlab = "Predicted", ylab = "Actual")
abline(0, 1) 

#rmse(worldmap_lm_pred, worldmap_b_no_geometry_tst$cumconfirmed) #279547822
#We did not get lower test RMSE. The most obvious linear regression couldn't beat the tree! 
```
The tree after pruning seems to be easier to interpret and has a better graphical
representation.


```{r, echo = FALSE, warning=FALSE}
# Boosting
require(gbm)
set.seed(18)
boost.worldmap_b_no_geometry=gbm(cumconfirmed~area+pop_est+pop_est_dens+economy+cumdeaths+cumrecovered,data=worldmap_b_no_geometry_trn,distribution="gaussian",n.trees=5000,interaction.depth=4)

summary(boost.worldmap_b_no_geometry)

par(mfrow=c(1,2)) 
plot(boost.worldmap_b_no_geometry,i="pop_est_dens", type = "l")
plot(boost.worldmap_b_no_geometry,i="economy", type = "l")
plot(boost.worldmap_b_no_geometry,i="cumrecovered", type = "l")
plot(boost.worldmap_b_no_geometry,i="cumdeaths", type = "l")
plot(boost.worldmap_b_no_geometry,i="pop_est", type = "l")

```
The best predictor seems to be economy.

cumrecovered estimate is the next best predictor used followed by cumdeaths, pop_est_dens, area and pop_est.

The tree() function has used pop_est and economy. for building the Regression tree.

